Nonparametric Goodness-of-Fit Tests for Discrete Null Distributions

### 2011-08-25 rjournal/AE.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@73 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

For larger sample sizes, the classical Kolmogorov-Smirnov test is used
and is known to produce conservative p-values for discrete distributions;
the revised \code{ks.test()} supports estimation
of p-values via simulation if desired.

@@

@@

data:  c(0, 1) 
D = 0.5, $p$-value = 0.5
alternative hypothesis: two-sided 
\end{Soutput}
\end{Schunk}
Instead, the value of $D$ given in equation (1)
should be 0 and the associated $p$-value should be 1.  
Our revision of \code{ks.test()}
fixes this problem when the user provides a discrete distribution:
\begin{Schunk} 
\begin{Sinput}
> library(ks.test)
> ks.test::ks.test(c(0, 1), ecdf(c(0, 1)))
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

@@

Next, we simulate a sample of size 25 from the discrete uniform distribution on
the integers $\{1, 2, \ldots, 10\}$ and show usage of the
new \code{ks.test()} implementation.  The first is the default two-sided test,
where the exact $p$-value is obtained using the methods of \citet{gleser85}.
\begin{Schunk}
\begin{Sinput}
> set.seed(1)
> x <- sample(1:10, 25, replace = TRUE)
> x
\end{Sinput}
\begin{Soutput}
 [1]  3  4  6 10  3  9 10  7  7  1  3  2  7
[14]  4  8  5  8 10  4  8 10  3  7  2  3
\end{Soutput}
\begin{Sinput}
> ks.test::ks.test(x, ecdf(1:10))
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

data:  x 
D = 0.08, $p$-value = 0.9354
alternative hypothesis: two-sided 
\end{Soutput}
\end{Schunk}
Next, we conduct the default one-sided test, where Conover's method
provides the exact $p$-value (up to the numerical precision of the
implementation):
\begin{Schunk}
\begin{Sinput}
> ks.test::ks.test(x, ecdf(1:10),
+   alternative = "g")
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

data:  x 
D^+ = 0.04, $p$-value = 0.7731
alternative hypothesis:
the CDF of x lies above the null hypothesis 
\end{Soutput}
\end{Schunk}
In contrast, the option \code{exact=FALSE} results in the
$p$-value obtained by applying the classical Kolmogorov-Smirnov
test, resulting in a conservative $p$-value:
\begin{Schunk}
\begin{Sinput}
> ks.test::ks.test(x, ecdf(1:10),
+   alternative = "g", exact = FALSE)
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

data:  x 
D^+ = 0.04, $p$-value = 0.9231
alternative hypothesis:
the CDF of x lies above the null hypothesis 
\end{Soutput}
\end{Schunk}
The $p$-value may also be estimated via a Monte Carlo simulation:
\begin{Schunk}
\begin{Sinput}
> ks.test::ks.test(x, ecdf(1:10),
+   alternative = "g",
+   simulate.p.value=TRUE, B=10000)
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

@@

A different toy example shows the dangers of using R's existing
\code{ks.test()} function with discrete data:
\begin{Schunk}
\begin{Sinput}
> ks.test::ks.test(rep(1, 3), ecdf(1:3))
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

@@

Further 
generalizations of goodness-of-fit tests for discrete distributions
are described in the extended study of \cite{dewev1973}.  There are
existing R packages for certain type of Cram\'{e}r-von Mises 
goodness-of-fit tests for continuous distributions.
Functions implemented in package \pkg{nortest} 
\citep{nortest} focus on the
composite hypothesis of normality, while
package \pkg{ADGofTest} \citep{ADGofTest} provides the Anderson-Darling variant
of the test for general continuous distributions.  Packages
\pkg{CvM2SL1Test} \citep{CvM2SL1Test} and \pkg{CvM2SL2Test} 
\citep{CvM2SL2Test} provide
two-sample Cram\'{e}r-von Mises tests with continuous distributions.
Package \pkg{cramer} \citep{cramer} offers a multivariate Cram\'{e}r test for
the two-sample problem.

>>>>

For larger sample sizes (or when requested for smaller sample sizes),
the classical Kolmogorov-Smirnov test is used
and is known to produce conservative p-values for discrete distributions;
the revised \code{ks.test()} supports estimation
of p-values via simulation if desired.

@@

NOTE: NEED TO ADD SIMUALTED P-VALUES.

@@

data:  c(0, 1) 
D = 0.5, $p$-value = 0.5
alternative hypothesis: two-sided 
\end{Soutput}
\end{Schunk}
Instead, the value of $D$ given in equation (1)
should be 0 and the associated $p$-value should be 1.  
Our revision of \code{ks.test()}
fixes this problem when the user provides a discrete distribution:
\begin{Schunk} 
\begin{Sinput}
> library(ks.test)
> dgof::ks.test(c(0, 1), ecdf(c(0, 1)))
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

@@

Next, we simulate a sample of size 25 from the discrete uniform distribution on
the integers $\{1, 2, \ldots, 10\}$ and show usage of the
new \code{ks.test()} implementation.  The first is the default two-sided test,
where the exact $p$-value is obtained using the methods of \citet{gleser85}.
\begin{Schunk}
\begin{Sinput}
> set.seed(1)
> x <- sample(1:10, 25, replace = TRUE)
> x
\end{Sinput}
\begin{Soutput}
 [1]  3  4  6 10  3  9 10  7  7  1  3  2  7
[14]  4  8  5  8 10  4  8 10  3  7  2  3
\end{Soutput}
\begin{Sinput}
> dgof::ks.test(x, ecdf(1:10))
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

data:  x 
D = 0.08, $p$-value = 0.9354
alternative hypothesis: two-sided 
\end{Soutput}
\end{Schunk}
Next, we conduct the default one-sided test, where Conover's method
provides the exact $p$-value (up to the numerical precision of the
implementation):
\begin{Schunk}
\begin{Sinput}
> dgof::ks.test(x, ecdf(1:10),
+   alternative = "g")
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

data:  x 
D^+ = 0.04, $p$-value = 0.7731
alternative hypothesis:
the CDF of x lies above the null hypothesis 
\end{Soutput}
\end{Schunk}
In contrast, the option \code{exact=FALSE} results in the
$p$-value obtained by applying the classical Kolmogorov-Smirnov
test, resulting in a conservative $p$-value:
\begin{Schunk}
\begin{Sinput}
> dgof::ks.test(x, ecdf(1:10),
+   alternative = "g", exact = FALSE)
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

data:  x 
D^+ = 0.04, $p$-value = 0.9231
alternative hypothesis:
the CDF of x lies above the null hypothesis 
\end{Soutput}
\end{Schunk}
The $p$-value may also be estimated via a Monte Carlo simulation:
\begin{Schunk}
\begin{Sinput}
> dgof::ks.test(x, ecdf(1:10),
+   alternative = "g",
+   simulate.p.value=TRUE, B=10000)
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

@@

A different toy example shows the dangers of using R's existing
\code{ks.test()} function with discrete data:
\begin{Schunk}
\begin{Sinput}
> dgof::ks.test(rep(1, 3), ecdf(1:3))
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

@@

Further 
generalizations of goodness-of-fit tests for discrete distributions
are described in the extended study of \cite{dewev1973}.  There are
existing R packages for certain type of Cram\'{e}r-von Mises 
goodness-of-fit tests for continuous distributions.
Functions implemented in package \pkg{nortest} 
\citep{nortest} focus on the
composite hypothesis of normality, while
package \pkg{ADGofTest} \citep{ADGofTest} provides the Anderson-Darling variant
of the test for general continuous distributions.  Packages
\pkg{CvM2SL1Test} \citep{CvM2SL1Test} and \pkg{CvM2SL2Test} 
\citep{CvM2SL2Test} provide
two-sample Cram\'{e}r-von Mises tests with continuous distributions.
Package \pkg{cramer} \citep{cramer} offers a multivariate Cram\'{e}r test for
the two-sample problem.  Finally, we note that the discrete goodness-of-fit tests
discussed in this paper do not allow the estimation of parameters of the
hypothesized null distributions (see \cite{LSS2007} for example).
